---
title: "Project testing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question: Does variation in karyotype accelerate the rate of change of Genome size?

```{r}
library(pdftools)
library(ggplot2)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ape)
library(phangorn)
library(phytools)
library(adephylo)
```
# Data wrangling
```{r}
#All to get data from pdf to r
# download pdf
raw_text <-purrr::map("41559_2018_674_MOESM1_ESM.pdf", pdftools::pdf_text)

raw_text <- purrr::map(raw_text, ~ str_split(.x, "\\n") %>% unlist())
  # Concatenate the split pages
  raw_text <- reduce(raw_text, c)
  raw2 <- as.data.frame(raw_text)
#remove back-to-back " 's"
sum <- 0
 repeat {
     raw2$raw_text <- gsub('            ', ' ', raw2$raw_text)
     raw2$raw_text <- gsub('    ', ' ', raw2$raw_text)
     raw2$raw_text <- gsub('  ', ' ', raw2$raw_text)
     sum <- sum + 1
if(sum == 3) {
break
}

 }
raw2$raw_text <- gsub('^ ', '', raw2$raw_text)

#Separate each statistic into its own column
  rawX <- separate(raw2, col=raw_text, into=c("Genus", "Species", "Order", "Family", "n", "Mean", "Median", "SD", "Min.", "Max.","Source"), sep = " ")
#Remove now jumbled header 
rawX <- rawX[-c(1:9), ]
#Remove blank rows
cVal<- rawX[!is.na(rawX$Order), ]
#combine species and genus
cVal$Species <- paste0(cVal$Genus, " ", cVal$Species)
hist(as.numeric(cVal$Mean), breaks = 50, xlab = "C-value", col = "orange", main = "Histogram of C-value")
hist(log(as.numeric(cVal$Mean)), breaks = 50, xlab = "Log(C-value)", col = "purple", main = "Histogram of Log(C-value)")
cVal$logC <- log(as.numeric(cVal$Mean))
#It might be worth log-transforming these values
```

```{r}
karyo <- read.csv("amphib_2022_10_29.csv")
dim(karyo)
#Remove duplicate entries for same species
#   -note: this could be done more empirically (should make sure none of the duplicates conflict)
karyo <- karyo[!duplicated(karyo$Species), ]
#I know there's a better way, or at least a way to check if two of the same species have different karyotypes reported: karyo[duplicated(karyo$Species), "Species"]

#Merge dataframes, keeping only shared taxa
joint <- dplyr::inner_join(cVal, karyo, by = c("Species", "Order", "Family"))
joint <- joint[, c("Species", "Genus", "Family", "Order", "Mean", "Median", "Diploid.number", "Fundamental.number", "logC")]
```

```{r}
#joint <- polyploid species
#Reading in data from Hime et al
#For Future, load in phylogenies first, then use tipnode labels directly to select species represented in both phylogenies
taxa <- read.csv("taxa.csv")
taxa$Species <- paste0(taxa$Genus, " ", taxa$Species)
#Merge dataframes, keeping only shared taxa
phyloDat <- dplyr::inner_join(joint, taxa, by = c("Species"))
#Remove duplicated columns
phyloDat <- phyloDat[, !(colnames(phyloDat) %in% c("Genus.y", "Genus.x", "Family.x", "Order.x", "AHE_ID", "Institution", "Order.y", "Family.y", "ID"))]
phyloDat <- phyloDat[-50, ]
```
# Loading in Phylogenies THIS CHUNK IS NOW OUT OF DATE BECAUSE IT DOESN'T YET SPLIT OFF CAUDATA
```{r fig.width=5, fig.height=10}
phyloDat$temp <- paste0(phyloDat$Species, "  ", phyloDat$Mean)
#Phylo from Pyron 2014
phy <- read.tree("amph_shl_dates_trim.tre")
phyloDat$Species <- gsub(" ", "_", phyloDat$Species)
#Drop all tips not represented in dataset
noKeep <- phy$tip.label[! phy$tip.label %in% phyloDat$Species]
phy_reduced = drop.tip(phy, noKeep)


#Load in more recent phylogeny 
phy2 <- read.tree("HimePhy.tre")
str(phy2)

#which tips are in phy reduced are in phy reduced 2
noKeep <- phy2$tip.label[! phy2$tip.label %in% phy_reduced$tip.label]
phy_reduced2 = drop.tip(phy2, noKeep)
#Drop the one tip not shared between phy reduced and phy reduced 2
phy_reduced <- drop.tip(phy_reduced, phy_reduced$tip.label[!phy_reduced$tip.label %in% phy_reduced2$tip.label])
str(phy_reduced)
#Directly comparable dataset created

#Make sure they're ultrametric
is.ultrametric(phy_reduced)
is.ultrametric(phy_reduced2)
#Should already be ultrametric, so force R to recognize that
phy_reduced2 <- force.ultrametric(phy_reduced2, method = "extend")

#create .tre files for newly reduced phylogenies
#write.tree(phy_reduced, file = "reduced.tre")
#write.tree(phy_reduced2, file = "reduced2.tre")

plot(phy_reduced)
# Now to check min branch length:
min(phy_reduced$edge.length)

plot(phy_reduced2)
is.ultrametric(phy_reduced2)
min(phy_reduced2$edge.length)

```

```{r}
#Remove tips not included in reduced phylogeny
amphib <- phyloDat[phyloDat$Species %in% phy_reduced$tip.label, ] 
amphib$logC <- paste0(amphib$Species, "  ", amphib$logC)
#write.table(amphib$temp, file = "amphib$temp.txt")
#write.table(amphib$logC, file = "amphib$log.txt")
```
# BAMM
```{r}
#Setting up BAMM
library(BAMMtools)
#setBAMMpriors(phy = phy_reduced, traits = "amphib$temp.txt", outfile = "pyronPrior.txt")
#setBAMMpriors(phy = phy_reduced, traits = "amphib$log.txt", outfile = "logPyronPrior.txt")

#set up BAMM priors
#setBAMMpriors(phy = phy_reduced2, traits = "amphib$temp.txt", outfile = "HimePrior.txt")
#setBAMMpriors(phy = phy_reduced2, traits = "amphib$log.txt", outfile = "logHimePrior.txt")
```

```{r fig.width=5, fig.height=16}
#Full run w/ both phylogenies
joint$Species <- gsub(" ", "_", joint$Species)
#Drop all tips not represented in dataset
noKeep <- phy$tip.label[! phy$tip.label %in% joint$Species]
full = drop.tip(phy, noKeep)
#distRoot(full)
noKeep <- phy_reduced2$tip.label[! phy_reduced2$tip.label %in% joint$Species]
full2 = drop.tip(phy_reduced2, noKeep)
is.ultrametric(full)
is.ultrametric(full2)
full <- force.ultrametric(full, method = "extend")
full2 <- force.ultrametric(full2, method = "extend")
#Convert branch length to millions of years for Hime phylogeny
full2$edge.length <- full2$edge.length*1000.1525656456
#distRoot(full2)



#Files created for BAMM before I realized I needed to analyze Caudata seperately to account for jump on caudatan stem
#write.table(sub$Cbamm, file = "sub$Cbamm.txt")
#write.table(sub2$Cbamm, file = "sub2$Cbamm.txt")
#write.tree(full, file = "pyron.tre")
#write.tree(full2, file = "hime.tre")

#setBAMMpriors(phy = full, traits = "sub$Cbamm.txt", outfile = "FullpyronPrior.txt")
#setBAMMpriors(phy = full2, traits = "sub2$Cbamm.txt", outfile = "FullHimePrior.txt")

#Building out analysis for Caudata and others separate
sally <- getMRCA(full, c("Andrias_japonicus", "Plethodon_jordani"))
yes <- getDescendants(full, sally)
sallys <- keep.tip(full, yes)
#plot(sallys)
noSally <- drop.tip(full, yes)
write.tree(sallys, file = "pyron_mander.tre")
write.tree(noSally, file = "pyron_nomander.tre")

sally <- getMRCA(full2, c("Andrias_japonicus", "Plethodon_jordani"))
yes <- getDescendants(full2, sally)
sallys2 <- keep.tip(full2, yes)
noSally2 <- drop.tip(full2, yes)
write.tree(sallys2, file = "hime_mander.tre")
write.tree(noSally2, file = "hime_nomander.tre")

#Create files needed for BAMM
joint$Cbamm <- paste0(joint$Species, "  ", joint$logC)
subMan <- joint[joint$Species %in% sallys$tip.label, ]
subNo <- joint[joint$Species %in% noSally$tip.label, ]
subMan2 <- joint[joint$Species %in% sallys2$tip.label, ]
subNo2 <- joint[joint$Species %in% noSally2$tip.label, ]

#write.table(subMan$Cbamm, file = "subMan$Cbamm.txt")
#write.table(subNo$Cbamm, file = "subNo$Cbamm.txt")
#write.table(subMan2$Cbamm, file = "subMan2$Cbamm.txt")
#write.table(subNo2$Cbamm, file = "subNo2$Cbamm.txt")

setBAMMpriors(phy = sallys, traits = "subMan$Cbamm.txt", outfile = "FullpyronPrior_mander.txt")
setBAMMpriors(phy = noSally, traits = "subNo$Cbamm.txt", outfile = "FullpyronPrior_nomander.txt")

setBAMMpriors(phy = sallys2, traits = "subMan2$Cbamm.txt", outfile = "FullHimePrior_mander.txt")
setBAMMpriors(phy = noSally2, traits = "subNo2$Cbamm.txt", outfile = "FullHimePrior_nomander.txt")
```

# BAMM results for all individuals for which I have both Karyotype and C-value
```{r}
setwd("/Users/raymondfedrick/Desktop/Final_Phylo/MCMCex")

## Syntax used throughout code below
#    -P = Pyron Phylogeny without salamanders
#    -PM = Pyron Phylogeny only including salamanders
#    -H = Hime Phylogeny without salamanders
#    -HM = Hime Phylogeny only including salamanders

mcmcoutP <- read.csv("mcmc_outP.txt", header=T)
plot(mcmcoutP$logLik ~ mcmcoutP$generation)
mcmcoutPM <- read.csv("mcmc_outPM.txt", header=T)
plot(mcmcoutPM$logLik ~ mcmcoutPM$generation)

mcmcoutH <- read.csv("mcmc_outH.txt", header=T)
plot(mcmcoutH$logLik ~ mcmcoutH$generation)
mcmcoutHM <- read.csv("mcmc_outHM.txt", header=T)
plot(mcmcoutHM$logLik ~ mcmcoutHM$generation)

burnstartP <- floor(0.25 * nrow(mcmcoutP))
postburnP <- mcmcoutP[burnstartP:nrow(mcmcoutP), ]
burnstartPM <- floor(0.25 * nrow(mcmcoutPM))
postburnPM <- mcmcoutPM[burnstartPM:nrow(mcmcoutPM), ]

burnstartH <- floor(0.25 * nrow(mcmcoutH))
postburnH <- mcmcoutH[burnstartH:nrow(mcmcoutH), ]
burnstartHM <- floor(0.25 * nrow(mcmcoutHM))
postburnHM <- mcmcoutHM[burnstartHM:nrow(mcmcoutHM), ]

post_probsP <- table(postburnP$N_shifts) / nrow(postburnP)
names(post_probsP)
plotPrior("mcmc_outP.txt", expectedNumberOfShifts = 1)
post_probsPM <- table(postburnPM$N_shifts) / nrow(postburnPM)
names(post_probsPM)
plotPrior("mcmc_outPM.txt", expectedNumberOfShifts = 1)

post_probsH <- table(postburnH$N_shifts) / nrow(postburnH)
names(post_probsH)
plotPrior("mcmc_outH.txt", expectedNumberOfShifts = 1)
post_probsHM <- table(postburnHM$N_shifts) / nrow(postburnHM)
names(post_probsHM)
plotPrior("mcmc_outHM.txt", expectedNumberOfShifts = 1)
```

```{r fig.width=15, fig.height=15}
setwd("/Users/raymondfedrick/Desktop/Final_Phylo/MCMCex")
#computeBayesFactors("event_dataP.txt", expectedNumberOfShifts=1, burnin=0.25)
bammedP <- getEventData(noSally, "event_dataP.txt", burnin = 0.25, type = "trait")
plot.bammdata(bammedP, lwd=2, labels = TRUE, legend = TRUE, method = "polar")
bammedPM <- getEventData(sallys, "event_dataPM.txt", burnin = 0.25, type = "trait")
plot.bammdata(bammedPM, lwd=2, labels = TRUE, legend = TRUE, method = "polar")

bammedH <- getEventData(noSally2, "event_dataH.txt", burnin = 0.25, type = "trait")
plot.bammdata(bammedH, lwd=2, labels = TRUE, legend = TRUE)
bammedHM <- getEventData(sallys2, "event_dataHM.txt", burnin = 0.25, type = "trait")
plot.bammdata(bammedHM, lwd=2, labels = TRUE, legend = TRUE)
```

```{r, fig.width=8, fig.height=8}
#Data for rates in individual branches
allratesP <- getMeanBranchLengthTree(bammedP, rate = "trait")
allratesPM <- getMeanBranchLengthTree(bammedPM, rate = "trait")

allratesH <- getMeanBranchLengthTree(bammedH, rate = "trait")
allratesHM <- getMeanBranchLengthTree(bammedHM, rate = "trait")

#Characterizing rate data
hist(allratesP$phy$edge.length, breaks = 50, main = "Pyron rates without Salamanders", xlab = "C- value evolutionary rate", col = "dark green")
hist(allratesPM$phy$edge.length, breaks = 50, main = "Pyron rates with Salamanders", xlab = "C- value evolutionary rate", col = "orange")
hist(allratesH$phy$edge.length, breaks = 50, main = "Hime rates without Salamanders", xlab = "C- value evolutionary rate", col = "brown")
hist(allratesHM$phy$edge.length, breaks = 50, main = "Hime rates with Salamanders", xlab = "C- value evolutionary rate", col = "yellow")

plot(allratesH$phy)
plot(allratesHM$phy)
```

## Identifying Branches with predicted karyotype change
```{r, fig.width=8, fig.height=8, warning = FALSE}
#Hime phylo, Salamanders only
#For edges connecting to tip nodes
nums <- match(c("Andrias_japonicus", "Siren_intermedia", "Notophthalmus_viridescens", "Rhyacotriton_olympicus", "Nyctanolis_pernix"), allratesHM$phy$tip.label)
kEdgesHM <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgesHM[i] <- allratesHM$phy$edge.length[allratesHM$phy$edge[,2] == nums[i]]
}
#For edges one node deeper than the tip
nums <- match("Necturus_maculosus", allratesHM$phy$tip.label)
kEdgeHMtemp <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgeHMtemp[i] <- allratesHM$phy$edge[allratesHM$phy$edge[,2] == nums[i]]
}
for (i in 1:length(nums)) {
  kEdgeHMtemp[i] <- allratesHM$phy$edge.length[allratesHM$phy$edge[,2] == kEdgeHMtemp[i]]
}
kEdgesHM <- c(kEdgesHM, kEdgeHMtemp)
#Edges two nodes deeper than the tip
nums <- match("Salamandrina_terdigitata", allratesHM$phy$tip.label)
kEdgeHMtemp <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgeHMtemp[i] <- allratesHM$phy$edge[allratesHM$phy$edge[,2] == nums[i]]
}
for (i in 1:length(nums)) {
  kEdgeHMtemp[i] <- allratesHM$phy$edge[allratesHM$phy$edge[,2] == kEdgeHMtemp[i]]
}
kEdgeHMtemp
for (i in 1:length(nums)) {
  kEdgeHMtemp[i] <- allratesHM$phy$edge.length[allratesHM$phy$edge[,2] == kEdgeHMtemp[i]]
}
kEdgesHM <- c(kEdgesHM, kEdgeHMtemp)

noEdgesHM <- allratesHM$phy$edge.length[!allratesHM$phy$edge.length %in% kEdgesHM]

#Hime phylogeny, without salamanders
#For edges connecting to tip nodes
nums <- match(c("Cardioglossa_leucomystax", "Astylosternus_diadematus", "Microhyla_ornata", "Plethodontohyla_notosticta", "Sooglossus_sechellensis", "Odontophrynus_occidentalis", "Pleurodema_bibroni", "Batrachyla_taeniata", "Limnodynastes_dumerilii", "Hymenochirus_boettgeri", "Rhinophrynus_dorsalis", "Alytes_obstetricans", "Geotrypetes_seraphini", "Grandisonia_alternans", "Boulengerula_taitana"), allratesH$phy$tip.label)
kEdgesH <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgesH[i] <- allratesH$phy$edge.length[allratesH$phy$edge[,2] == nums[i]]
}
#For edges one node deeper than the tip
nums <- match(c("Pyxicephalus_adspersus", "Dermatonotus_muelleri", "Spea_hammondii", "Pipa_pipa", "Discoglossus_pictus", "Leiopelma_hochstetteri", "Gymnopis_multiplicata", "Boulengerula_taitana"), allratesH$phy$tip.label)
kEdgeHtemp <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgeHtemp[i] <- allratesH$phy$edge[allratesH$phy$edge[,2] == nums[i]]
}
for (i in 1:length(nums)) {
  kEdgeHtemp[i] <- allratesH$phy$edge.length[allratesH$phy$edge[,2] == kEdgeHtemp[i]]
}
kEdgesH <- c(kEdgesH, kEdgeHtemp)

noEdgesH <- allratesH$phy$edge.length[!allratesH$phy$edge.length %in% kEdgesH]

#Pyron No Manders
#For edges connecting to tip nodes
nums <- match(c("Dendrobates_tinctorius", "Allophryne_ruthveni", "Adenomera_hylaedactyla", "Pseudacris_brimleyi", "Dendropsophus_nanus", "Batrachyla_taeniata", "Zachaenus_parvulus", "Ceratophrys_ornata", "Eleutherodactylus_planirostris", "Eleutherodactylus_heminota", "Eleutherodactylus_abbotti", "Eleutherodactylus_martinicensis", "Eleutherodactylus_richmondi", "Mantidactylus_biporus", "Phrynobatrachus_natalensis", "Astylosternus_diadematus", "Xenopus_vestitus", "Xenopus_ruwenzoriensis", "Hymenochirus_boettgeri", "Pipa_pipa", "Rhinophrynus_dorsalis", "Ascaphus_truei", "Gegeneophis_ramaswamii", "Schistometopum_gregorii", "Geotrypetes_seraphini", "Chthonerpeton_indistinctum", "Boulengerula_taitana"), allratesP$phy$tip.label)
kEdgesP <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgesP[i] <- allratesP$phy$edge.length[allratesP$phy$edge[,2] == nums[i]]
}
#For edges one node deeper than the tip
nums <- match(c("Nimbaphrynoides_occidentalis", "Acris_crepitans", "Litoria_bicolor", "Limnodynastes_dumerilii", "Mantidactylus_curtus", "Rana_chensinensis", "Gastrophryne_carolinensis", "Hemisus_marmoratus", "Arthroleptis_stenodactylus", "Alytes_cisternasii", "Bombina_variegata", "Leiopelma_hochstetteri"), allratesP$phy$tip.label)
kEdgePtemp <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgePtemp[i] <- allratesP$phy$edge[allratesP$phy$edge[,2] == nums[i]]
}
for (i in 1:length(nums)) {
  kEdgePtemp[i] <- allratesP$phy$edge.length[allratesP$phy$edge[,2] == kEdgePtemp[i]]
}
kEdgesP <- c(kEdgesP, kEdgePtemp)
#For edges two nodes deeper than the tip
nums <- match(c("Pseudopaludicola_mystacalis", "Proceratophrys_boiei", "Eleutherodactylus_richmondi", "Xenopus_ruwenzoriensis", "Discoglossus_jeanneae"), allratesP$phy$tip.label)
kEdgePtemp <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgePtemp[i] <- allratesP$phy$edge[allratesP$phy$edge[,2] == nums[i]]
}
for (i in 1:length(nums)) {
  kEdgePtemp[i] <- allratesP$phy$edge[allratesP$phy$edge[,2] == kEdgePtemp[i]]
}
for (i in 1:length(nums)) {
  kEdgePtemp[i] <- allratesP$phy$edge.length[allratesP$phy$edge[,2] == kEdgePtemp[i]]
}
kEdgesP <- c(kEdgesP, kEdgePtemp)
#For edges three nodes deeper than the tip
nums <- match(c("Crinia_signifera"), allratesP$phy$tip.label)
kEdgePtemp <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgePtemp[i] <- allratesP$phy$edge[allratesP$phy$edge[,2] == nums[i]]
}
for (i in 1:length(nums)) {
  kEdgePtemp[i] <- allratesP$phy$edge[allratesP$phy$edge[,2] == kEdgePtemp[i]]
}
for (i in 1:length(nums)) {
  kEdgePtemp[i] <- allratesP$phy$edge[allratesP$phy$edge[,2] == kEdgePtemp[i]]
}
for (i in 1:length(nums)) {
  kEdgePtemp[i] <- allratesP$phy$edge.length[allratesP$phy$edge[,2] == kEdgePtemp[i]]
}
kEdgesP <- c(kEdgesP, kEdgePtemp)

noEdgesP <- allratesP$phy$edge.length[!allratesP$phy$edge.length %in% kEdgesP]

#Pyron Salamanders only
#For edges connecting to tip nodes
nums <- match(c("Rhyacotriton_olympicus", "Ambystoma_texanum", "Ambystoma_laterale", "Hynobius_retardatus"), allratesPM$phy$tip.label)
kEdgesPM <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgesPM[i] <- allratesPM$phy$edge.length[allratesPM$phy$edge[,2] == nums[i]]
}
#For edges one node deeper than the tip
nums <- match(c("Batrachoseps_attenuatus", "Proteus_anguinus", "Taricha_rivularis", "Salamandrina_terdigitata", "Pseudobranchus_striatus", "Onychodactylus_fischeri", "Cryptobranchus_alleganiensis"), allratesPM$phy$tip.label)
kEdgePMtemp <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgePMtemp[i] <- allratesPM$phy$edge[allratesPM$phy$edge[,2] == nums[i]]
}
for (i in 1:length(nums)) {
  kEdgePMtemp[i] <- allratesPM$phy$edge.length[allratesPM$phy$edge[,2] == kEdgePMtemp[i]]
}
kEdgesPM <- c(kEdgesPM, kEdgePMtemp)
#For edges two nodes deeper than the tip
nums <- match(c("Cryptobranchus_alleganiensis"), allratesPM$phy$tip.label)
kEdgePMtemp <- rep(NA, length(nums))
for (i in 1:length(nums)) {
  kEdgePMtemp[i] <- allratesPM$phy$edge[allratesPM$phy$edge[,2] == nums[i]]
}
for (i in 1:length(nums)) {
  kEdgePMtemp[i] <- allratesPM$phy$edge[allratesPM$phy$edge[,2] == kEdgePMtemp[i]]
}
for (i in 1:length(nums)) {
  kEdgePMtemp[i] <- allratesPM$phy$edge.length[allratesPM$phy$edge[,2] == kEdgePMtemp[i]]
}
kEdgesPM <- c(kEdgesPM, kEdgePMtemp)

noEdgesPM <- allratesPM$phy$edge.length[!allratesPM$phy$edge.length %in% kEdgesPM]
```

## Plotting Genome size evolution rates by predicted karyotype change
```{r, fig.width=8, fig.height=8}

## Refresher on syntax used throughout code below
#    -P = Pyron Phylogeny without salamanders
#    -PM = Pyron Phylogeny only including salamanders
#    -H = Hime Phylogeny without salamanders
#    -HM = Hime Phylogeny only including salamanders

library(ggplot2)
dfPM <- data.frame(Karyotype = n<-c(rep("Karyo Change", length(kEdgesPM)), rep("No Change", length(noEdgesPM))), Rate = c(kEdgesPM, noEdgesPM))
ggplot(dfPM, aes(x = Karyotype, y = Rate)) +            
  geom_boxplot(fill = c("red", 4), color = "black") +
  ggtitle("C-Value Evolutionary rates by Karyotype change", subtitle = "Pyron Phylogeny only Salamanders")

dfP <- data.frame(Karyotype = n<-c(rep("Karyo Change", length(kEdgesP)), rep("No Change", length(noEdgesP))), Rate = c(kEdgesP, noEdgesP))
ggplot(dfP, aes(x = Karyotype, y = Rate)) +            
  geom_boxplot(fill = c("red", 4), color = "black") +
  ggtitle("C-Value Evolutionary rates by Karyotype change", subtitle = "Pyron Phylogeny without salamanders")

dfHM <- data.frame(Karyotype = n<-c(rep("Karyo Change", length(kEdgesHM)), rep("No Change", length(noEdgesHM))), Rate = c(kEdgesHM, noEdgesHM))
ggplot(dfHM, aes(x = Karyotype, y = Rate)) +            
  geom_boxplot(fill = c("red", 4), color = "black") +
  labs(title = "C-Value Evolutionary rates by Karyotype change", subtitle = "Hime Phylogeny only Salamanders")

dfH <- data.frame(Karyotype = n<-c(rep("Karyo Change", length(kEdgesH)), rep("No Change", length(noEdgesH))), Rate = c(kEdgesH, noEdgesH))
ggplot(dfH, aes(x = Karyotype, y = Rate)) +            
  geom_boxplot(fill = c("red", 4), color = "black") +
  labs(title = "C-Value Evolutionary rates by Karyotype change", subtitle = "Hime Phylogeny without salamanders")

dfPf <- rbind(dfP, dfPM)
ggplot(dfPf, aes(x = Karyotype, y = Rate)) +            
  geom_boxplot(fill = c("red", 4), color = "black") +
  ggtitle("C-Value Evolutionary rates by Karyotype change in Full Pyron Phylogeny")

dfHf <- rbind(dfH, dfHM)
ggplot(dfHf, aes(x = Karyotype, y = Rate)) +            
  geom_boxplot(fill = c("red", 4), color = "black") +
  labs(title = "C-Value Evolutionary rates by Karyotype change in Full Hime Phylogeny")


#NOW DO ANOVA
#It occurs to me that anova assumes normal distributions, and these rates really aren't normally distributed
#Kruskal Wallis test to correct for non-normality of the data
anPM <- aov(Rate ~ Karyotype, data = dfPM)
summary(anPM)
kruskal.test(Rate ~ Karyotype, data = dfPM)
anP <- aov(Rate ~ Karyotype, data = dfP)
summary(anP)
kruskal.test(Rate ~ Karyotype, data = dfP)
anHM <- aov(Rate ~ Karyotype, data = dfHM)
summary(anHM)
kruskal.test(Rate ~ Karyotype, data = dfHM)
anH <- aov(Rate ~ Karyotype, data = dfH)
summary(anH)
kruskal.test(Rate ~ Karyotype, data = dfH)

#Concatanating Salamander rates with rates from the rest of amphibia
anPf <- aov(Rate ~ Karyotype, data = dfPf)
summary(anPf)
#Additional non-parametric test to account for non-normality of the data
kruskal.test(Rate ~ Karyotype, data = dfPf)

anHf <- aov(Rate ~ Karyotype, data = dfHf)
summary(anHf)
#Additional non-parametric test to account for non-normality of the data
kruskal.test(Rate ~ Karyotype, data = dfHf)

#Mean of Pyron phylogeny with significant effect of karyotype change
mean(tempk <- c(kEdgesPM, kEdgesP))
mean(tempk <- c(noEdgesPM, noEdgesP))


#Median of Pyron phylogeny with significant effect of karyotype change
median(tempk <- c(kEdgesPM, kEdgesP))
median(tempk <- c(noEdgesPM, noEdgesP))

```





Real question: are the branches which are predicted to have experienced changes in karyotype evolving faster than those where karyotype is epected to have remained constant
```



#Code using MrBayes to reconstruct ancestral state
```{r, eval = FALSE}
#Create files needed for MRBASR
joint$Kar <- paste0(joint$Species, "  ", joint$Diploid.number)
himeKar <- joint[joint$Species %in% full2$tip.label, ]
pyronKar <- joint[joint$Species %in% full$tip.label, ]
sort(unique(himeKar$Diploid.number))
sort(unique(pyronKar$Diploid.number))
table(himeKar$Diploid.number)
table(pyronKar$Diploid.number)

#Limited to 6 sequential bins, so must group 10 and 26 karyotypes repsectively into 6 bins :()
#Hime
#0 - 18+22
#1 - 24
#2 -26
#3 - 28
#4 - 34+38
#5 - 46, 54, 60

#Pyron
#0- 14, 16, 18, 20, 22
#1 - 24
#2 -26
#3 - 28
#4 - 30, 32, 34, 36, 38, 40
#5 -42, 44, 46, 48, 52, 56, 60, 66, 72, 78, 108
#write.table(himeKar$Kar, file = "himeKar.txt")
#write.table(pyronKar$Kar, file = "pyronKar.txt")

# Assign this path and use it to set the R working directory
MBASR.directory="/Users/raymondfedrick/Desktop/MBASR"
setwd(MBASR.directory)


# Load this source file
source("MBASR.load.functions.R")

#Hime Run
# Set these file names
file.name.tree="hime.nwk"
file.name.trait.data="himeFake.txt"
file.name.plot.settings="plot.settings.1.txt"

# Set these analysis parameters
character.type="ordered"
n.samples=1000

# Run the analysis
MBASR(file.name.tree, file.name.trait.data, file.name.plot.settings, character.type, n.samples)

#Pyron Run
# Set these file names
file.name.tree="pyron.nwk"
file.name.trait.data="pyronFake.txt"
file.name.plot.settings="plot.settings.2.txt"

# Set these analysis parameters
character.type="ordered"
n.samples=1000

# Run the analysis
MBASR(file.name.tree, file.name.trait.data, file.name.plot.settings, character.type, n.samples)


# Optional -- Replot the ASR result, after editing the plot settings file
plot.tree.with.pie.charts.custom <- function(file.name.tree,file.name.trait.data,file.name.plot.settings) {

##########

file.name.asr="MrBayes.ASR.results.txt"

###

source(file.name.plot.settings)

my.tip.label.offset=my.tip.label.shift.right
my.tip.pie.offset=my.tip.pie.shift.right

###

suppressWarnings(suppressMessages(library(ape)))
library(ape)

suppressWarnings(suppressMessages(library(phytools)))
library(phytools)

###

tree=read.tree(file.name.tree)

n_tips=length(tree$tip.label)
n_nodes=tree$Nnode
root_node_number=n_tips+1
last_node=root_node_number+n_nodes-1
node_numbers=seq(from=root_node_number,to=last_node,by=1)

tree$node.label=node_numbers

###

BLs.orig=tree$edge.length
n.edges=length(BLs.orig)
mean.BL=mean(BLs.orig)

scaling.factor=1
if(mean.BL<0.1) { scaling.factor=100 }
if(mean.BL<0.01) { scaling.factor=1000 }
if(mean.BL<0.001) { scaling.factor=10000 }
if(mean.BL<0.0001) { scaling.factor=100000 }
if(mean.BL<0.00001) { scaling.factor=1000000 }
if(mean.BL<0.000001) { scaling.factor=10000000 }

BLs.scaled=BLs.orig*scaling.factor
tree$edge.length[1:n.edges]=BLs.scaled

###

ASR_table=read.table(file.name.asr,sep="\t")
ASR_table=as.matrix(ASR_table)
rownames(ASR_table)=ASR_table[,1]
ASR_table=ASR_table[,-1]
mode(ASR_table)="numeric"
colnames(ASR_table)=NULL

###

tip_data=read.table(file.name.trait.data,sep="\t")
tip_data=as.matrix(tip_data)
colnames(tip_data)=NULL
rownames(tip_data)=tip_data[,1]
tip_data=as.matrix(tip_data[,-1])
colnames(tip_data)="state"

xyz=strsplit(tip_data[,1],"")
xyz=unlist(xyz)
options(warn=-1)
mode(xyz)="numeric"
options(warn=0)
max_state=max(xyz,na.rm=T)

first_state=0
all_states=seq(from=first_state,to=max_state,by=1)
n_states=length(all_states)

needed_rows=dim(tip_data)[1]
needed_cols=length(all_states)

state_names=c("14-22", "24", "26", "28", "30-40", "40+")
state_names=paste("Diploid number ",state_names,sep="")

###

c0="darkblue"
c1="royalblue1"
c2="turquoise"
c3="yellow"
c4="orange"
c5="red"
c6="purple"
c7="magenta"
c8="gray60"
c9="gray90"
my.colors=c(c0,c1,c2,c3,c4,c5,c6,c7,c8,c9)

my.colors=my.colors[1:n_states]

##########

tips_matrix=matrix(0,nrow=needed_rows,ncol=needed_cols)
rownames(tips_matrix)=rownames(tip_data)

current_taxon=1
repeat {
current_score=tip_data[current_taxon,1]
test_and=grep("&",current_score)

if(length(test_and)>0) {
current_score=strsplit(current_score,"&")
current_score=unlist(current_score)
options(warn=-1)
mode(current_score)="numeric"
options(warn=0)
destination_columns=current_score+1
current_strength=1/length(current_score)
round(current_strength,digits=3)
tips_matrix[current_taxon,destination_columns]=current_strength
}

if(length(test_and)==0) {
options(warn=-1)
mode(current_score)="numeric"
options(warn=0)
destination_columns=current_score+1
current_strength=1
tips_matrix[current_taxon,destination_columns]=current_strength
}

current_taxon=current_taxon+1
if(current_taxon==needed_rows+1) break }

###

Qs=which(tip_data=="?")
if(length(Qs)>0) {
tips_matrix[Qs,1:needed_cols]="?"
}

#tips_matrix

##########

tips_list_correct_order=tree$tip.label
tips_list_incorrect_order=rownames(tips_matrix)
my_match=match(tips_list_incorrect_order,tips_list_correct_order)

temp_matrix=cbind(tips_matrix,tips_list_correct_order)
temp_dim=dim(temp_matrix)
temp_cols=temp_dim[2]

abc=colnames(temp_matrix)
abc[temp_cols]="correct_seq"
colnames(temp_matrix)=abc

temp_matrix[,temp_cols]=my_match
options(warn=-1)
mode(temp_matrix)="numeric"
options(warn=0)

temp_matrix=as.matrix(temp_matrix[order(temp_matrix[,temp_cols]),])

dim_again=dim(tips_matrix)
cols_again=dim_again[2]
rows_again=dim_again[1]

tips_matrix[1:rows_again,1:cols_again]=temp_matrix[1:rows_again,1:cols_again]
options(warn=-1)
mode(tips_matrix)="numeric"
options(warn=0)

###

pdf(file="tree.plot.pdf",width=my.plot.width,height=my.plot.height)

if(show.node.numbers==FALSE) {
plot.phylo(tree,direction="rightwards",no.margin=TRUE,cex=my.tip.label.font.size,label.offset=my.tip.label.offset,edge.width=my.tree.edge.width,underscore=T)
}
tt
if(show.node.numbers==TRUE) {
plot.phylo(tree,direction="rightwards",no.margin=TRUE,cex=my.tip.label.font.size,label.offset=my.tip.label.offset,edge.width=my.tree.edge.width,underscore=T,show.node.label=T)
}

par(lwd=my.node.pie.border.width)
nodelabels(pie=ASR_table,cex=my.node.pie.size,piecol=my.colors)

par(lwd=my.tip.pie.border.width)
tiplabels(pie=tips_matrix,cex=my.tip.pie.size,adj=my.tip.pie.offset,piecol=my.colors)

legend(x=0,y=my.legend.shift.up,legend=state_names,box.col="white",title="Legend:", title.adj = -.05,cex=my.legend.size,fill=my.colors,horiz=T)

#

if(show.time.scale==TRUE) {

all.dist.nodes=dist.nodes(tree)
root.dist.nodes=all.dist.nodes[,root_node_number]
root.tips.dists=root.dist.nodes[1:n_tips]
root.age=max(root.tips.dists)

my.time.scale.n.ticks=10

one.tick.size=root.age/(my.time.scale.n.ticks-1)
scale.seq=seq(from=0,to=root.age,by=one.tick.size)

scale.seq.2=pretty(scale.seq,n=my.time.scale.n.ticks)
new.seq.length=length(scale.seq.2)
scale.seq.2[new.seq.length]=root.age

tick.labels=rev(scale.seq.2)
short.int=tick.labels[1]-tick.labels[2]
norm.int=tick.labels[2]-tick.labels[3]
adj.val=norm.int-short.int

scale.seq.3=scale.seq.2-adj.val
scale.seq.3[1]=scale.seq.2[1]
scale.seq.3[new.seq.length]=scale.seq.2[new.seq.length]

lastPP <- get("last_plot.phylo", envir = .PlotPhyloEnv)
xscale <- range(lastPP$xx)
tscale <- c(0,root.age)
beta <- diff(xscale)/diff(tscale)
alpha <- xscale[1] - beta * tscale[1]
ticks.at=beta*scale.seq.3+alpha

tick.labels[1]=NA

tick.labels=tick.labels+my.time.scale.youngest.tip
#tick.labels=round(tick.labels,digits=my.time.scale.label.n.digits)

tick.labels=tick.labels/scaling.factor

starting.options=options()
starting.scipen=starting.options$scipen
options(scipen=999)

axis(side=1,pos=my.time.scale.bar.shift.up,at=ticks.at,labels=tick.labels,cex.axis=my.tip.label.font.size,lwd=my.tree.edge.width,padj=-1*my.time.scale.label.shift.up,tcl=-0.1)

options(scipen=starting.scipen)

}
#

if(show.edge.transitions==TRUE) {

multiplier=state.decision.multiplier
file.name.asr.results="MrBayes.ASR.results.txt"
tips.nodes.table=make.tips.nodes.table(file.name.trait.data,file.name.tree,file.name.asr.results,multiplier)
edge.transition.table=make.edge.transition.table(file.name.tree,tips.nodes.table)
edge.transition.table[,4]=edge.transition.table[,4]+my.time.scale.youngest.tip
edge.transition.table[,5]=edge.transition.table[,5]+my.time.scale.youngest.tip
my.edge.labels=edge.transition.table[,8]

my.edge.transitions.shift.up=my.edge.transitions.shift.up*-1
edgelabels(my.edge.labels,frame="none",cex=my.edge.transitions.font.size,adj=c(my.edge.transitions.shift.left,my.edge.transitions.shift.up),col="#888888")
}

#

dev.off()

##########

msg="Tree plot was written to file."
msg=noquote(msg)

###

return(msg) }

#Create customized tree
plot.tree.with.pie.charts.custom(file.name.tree,file.name.trait.data,file.name.plot.settings)
```






